Toolset for an automated troubleshooting in SDNs

### 2013-11-05 draft.tex
flochecker completed

\maketitle
\begin{abstract}
In this paper, we propose a methodological review of the existing tools designed for Software-Defined networks troubleshooting. This review is organized using the SDN layering architecture, and aims to show the possibility of an automated localization of the bugs in the different layers of SDN. The toolset presented here allows administrators to more focus on fixing than localizing issues. We will go through the layers by choosing the most significant and relevant tools for each (pair of) layer(s), regarding other existing tools : NDB for spotting incoherences between policies and their effects, VeriFlow for verifying invariants in real-time, SOFT for spotting bugs in OpenFlow switches configuration, NICE either for the switch firmware of the controller software and FlowChecker supporting verification for different OpenFlow federated infrastructures.
\end{abstract}

@@

\section{SDN layering, the key used to an automated troubleshooting}\label{layering}
 % Expliquer ici la stack SDN (decomposition en layer) et donner l'intuition concernant la plus grande facilité pour remonter à l'origine d'un probleme.
Finding and solving network bugs are not the aim of SDN, but we can use it to re-think the way we troubleshoot networks.\\ The SDN architecture is decomposed into  layers, those layers can be represented in a two dimensionnal array. As you can see on Figure \ref{layers}, we have the two main layers called \textit{State layers} and \textit{Code layers}. The state layers hold a representation of the network's configuration for each parts of the network architecture. The code layers implement logic to maintain the mapping between two state layers. Each states layers should verify the equivalence properties, which means that each of them should correctly mapping every other state layer. The idea is that, for each policy, if the state layers are correctly mapped among each other, then the policy is  set and acts like it should.\\
On the Figure \ref{layers}, you have the following elements:\\
\textbf{Policy} - Policies are set up by the network administrator to configure the Logical View. These policies can be routing, access control or QoS policies. They are written inside a control application. See \cite{OpenStack_Quantum} for exemple.
\\
\textbf{Logical View} - Abstract representation of the network which aims to make things easier for the control application to create policies. A mapping between the Logical View and one or more Physical View are done by Network Hypervisor.\\
\textbf{Physical View} - This is a correspondance with the real network element, a representation of it handled by the network OS. The protocol used to configure the network element is one like OpenFlow \cite{OpenFlow}. A physical View has thus a one-to-one mapping with a real network element.
\\
\textbf{Device State} - State of the network element maintened by its firmware.
\\
\textbf{Hardware} - Network element. \\

@@

\begin{figure}
\centering
\includegraphics[scale=0.60]{layers.png}
\label{layers}
\caption{SDN architecture\label{layers}\protect\footnotemark}
\end{figure}
\footnotetext{Source : \cite{Heller:2013:LSL:2491185.2491197}}
\section{Network Troubleshooting - Tools}
\label{Network Troubleshooting - Tools}
\subsection{Network Debugger (NDB)}
SDNs bring a new way of building networks: like a programmer build a software, a network administrator will be able to program its network. The software programmer has tools to debug its program, like the well-known GDB.\\
NDB is a tool inspired from GDB, it aims at finding network error by identifying incorrect sequence of events. NDB has the ability to do a sort of \textit{breakpoint}, which is in fact a packet filter. When packet are matched by this filter, the breakpoint is triggered and the entire trace of that packet is displayed. The authors mention that the term \textit{tracepoint} should be more appropriate but the term \textit{breakpoint} is more familiar. NDB uses its \textit{backtrace} primitive to show the sequence of events for a packet. This trace is simply the sequence of forwarding actions applied on the packet by the switches through which it passed. For example of NDB utilisation, please refer to its paper \cite{ndb_paper}.\\
In our layering SDN architecture, NDB give us the opportunity to build a tool to confirm the right action of the policies implemented in the control application. For example, if a routing bug occurs in the network, a tool of top of NDB could observe paths taken by packets and complete switches states in order to conclude if the policies are correctly matched or not. If the policies are correctly matched, it means that the network administrator has written policies that dont't fit his needs. In this case, NDB has found the \textit{code layer} where the issue happen. Now if the policies are not correctly matched, then the issue does not come from an incorrect implementation. NDB has thus exclude the implementation to be the root of the bug and the search can continue among other layers, with other tools such as SOFT\cite{SOFT}.

@@

\subsection{SOFT - \textit{Systematic OpenFlow Testing}}
SOFT is a tool which aims to test the interoperability between OpenFlow switches. On our layering architecure, the OpenFlow protocol appears in the Physical View. SOFT has been designed to identify wrong behaviors of different switches running different implementation of the OpenFlow protocol. Thus, SOFT compares different Physical View of different switches. But SOFT can do more than that. It can also help to find inconsistency in the layers below the Physical Layers. SOFT can be used to check the consistency between the \textit{state layer} of Device State and the \textit{state layer} of Hardware. To see more details of how it works, refer to its paper \cite{SOFT}.

@@

The following tool is particularly adapted for verifying consistency between switches and controller across different OpenFlow federated infrastructures. FlowChecker, based on the ConfigChecker tool, provides a way to detect misconfigurations in the FlowTable of the switches\cite{canini2011automating}. For that goal, FlowChecker models the entire network as a state machine, which allows it to use the BDD-based (Binary Decision Diagram) model checker to verify network properties described by CTL (Computational Tree Logic) queries\cite{Al-Shaer:2010:FCA:1866898.1866905}.\\
To support inter-federated infrastructure, FlowChecker acts as a Master Controller that communicates with other controllers and switches. Those controllers, if interested to the experiment, send their FlowTable entries to the master controller, that verifies for inconsistencies, with a quadratic time function of the number of OpenFlow switches\cite{Al-Shaer:2010:FCA:1866898.1866905}.

>>>>

\maketitle
\begin{abstract}
In this paper, we propose a methodological review of the existing tools designed for Software-Defined networks troubleshooting. This review is organized using the SDN layering architecture, and aims to show the possibility of an automated localization of the bugs in the different layers of SDN. The toolset presented here allows administrators to more focus on fixing than localizing issues. We will go through the layers by choosing the most significant and relevant tools for each (pair of) layer(s) we are considering, regarding other existing tools : NDB for spotting incoherences between policies and their effects, VeriFlow for verifying invariants in real-time, SOFT for spotting bugs in OpenFlow switches configuration, NICE either for the switch firmware of the controller software and FlowChecker supporting verification for different OpenFlow federated infrastructures.
\end{abstract}

@@

\section{SDN layering, the key used to an automated troubleshooting}\label{layering}
 % Expliquer ici la stack SDN (decomposition en layer) et donner l'intuition concernant la plus grande facilité pour remonter à l'origine d'un probleme.
Finding and solving network bugs are not the aim of SDN, but we can use it to re-think the way we troubleshoot networks.\\ The SDN architecture is decomposed into  layers, those layers can be represented in a two dimensionnal array. As you can see on Figure \ref{layers}, we have the two main layers called \textit{State layers} and \textit{Code layers}. The state layers hold a representation of the network configuration for each parts of the network architecture. The code layers implement logic to maintain the mapping between two state layers. Each state layers should verify the equivalence properties, which means that each of them should correctly mapping every other state layer. The idea is that, for each policy, if the state layers are correctly mapped among each other, then the policy is  set and acts like it should.\\
On the Figure \ref{layers}, you have the following elements:\\
\textbf{Policy} - Policies are set up by the network administrator to configure the Logical View. These policies can be routing, access control or QoS policies. They are written inside a control application. See \cite{OpenStack_Quantum} for exemple.
\\
\textbf{Logical View} - Abstract representation of the network which aims to make things easier for the control application to create policies. A mapping between the Logical View and one or more Physical View are done by Network Hypervisor.\\
\textbf{Physical View} - This is a correspondance with the real network element, a representation of it handled by the network OS. The protocol used to configure the network element is one like OpenFlow \cite{OpenFlow}. A physical View has thus a one-to-one mapping with a real network element.
\\
\textbf{Device State} - State of the network element maintened by its firmware.
\\
\textbf{Hardware} - Network element. \\

@@

\begin{figure}
\centering
\includegraphics[scale=0.60]{layers.png}
\caption{SDN architecture\label{layers}\protect\footnotemark}
\end{figure}
\footnotetext{Source : \cite{Heller:2013:LSL:2491185.2491197}}
\section{Network Troubleshooting - Tools}
\label{Network Troubleshooting - Tools}
\subsection{Network Debugger (NDB)}
SDNs bring a new way of building networks: like a programmer build a software, a network administrator will be able to program its network. The software programmer has tools to debug its program, like the well-known GDB.\\
NDB is a tool inspired from GDB, it aims at finding network error by identifying incorrect sequence of events. NDB has the ability to do a sort of \textit{breakpoint}, which is in fact a packet filter. When packet are matched by this filter, the breakpoint is triggered and the entire trace of that packet is displayed. The authors mention that the term \textit{tracepoint} should be more appropriate but the term \textit{breakpoint} is more familiar. NDB uses its \textit{backtrace} primitive to show the sequence of events for a packet. This trace is simply the sequence of forwarding actions applied on the packet by the switches through which it passed. For example of NDB utilisation, please refer to its paper \cite{ndb_paper}.\\
In our layering SDN architecture, NDB give us the opportunity to build a tool to confirm the right action of the policies implemented in the control application. For example, if a routing bug occurs in the network, a tool of top of NDB could observe paths taken by packets and complete switches states in order to conclude if the policies are correctly matched or not. If the policies are correctly matched, it means that the network administrator has written policies that dont't fit his needs. In this case, NDB has found the \textit{code layer} where the issue happen. Now if the policies are not correctly matched, then the issue does not come from an incorrect implementation. NDB has thus exclude the implementation to be the root of the bug and the search can continue among other layers, with other tools such as SOFT\cite{SOFT}.

@@

\subsection{SOFT - \textit{Systematic OpenFlow Testing}}
SOFT is a tool which aims to test the interoperability between OpenFlow switches. On our layering architecure, the OpenFlow protocol appears in the Physical View. SOFT has been designed to identify wrong behaviors of different switches running different implementation of the OpenFlow protocol. Thus, SOFT compares different Physical View of different switches. But SOFT can do more than that. It can also help to find inconsistency in the layers below the Physical Layers. SOFT can be used to check the consistency between the \textit{state layer} of Device State and the \textit{state layer} of Hardware. To see more details of how it works, refer to its dedicated paper \cite{SOFT}.

@@

The following tool is particularly adapted for verifying consistency between switches and controller across different OpenFlow federated infrastructures. It operates in the the Physical View, just as SOFT does. FlowChecker, based on the ConfigChecker tool, provides a way to detect misconfigurations in the FlowTable of the switches\cite{canini2011automating}. For that goal, FlowChecker models the entire network as a state machine, which allows it to use the BDD-based (Binary Decision Diagram) model checker to verify network properties described by CTL (Computational Tree Logic) queries\cite{Al-Shaer:2010:FCA:1866898.1866905}.\\
To support inter-federated infrastructure, FlowChecker acts as a Master Controller that communicates with other controllers and switches. Those controllers, if interested to the experiment, send their FlowTable entries to the master controller, that verifies for inconsistencies, with a quadratic time function of the number of OpenFlow switches\cite{Al-Shaer:2010:FCA:1866898.1866905}.
